The Importance of Cell Connectivity for Efficient and Accurate Modeling of Flow through Dipping Aquifers with MODFLOW 6

### 2024-06-10 latex/article.tex
Add mfusg citation to section on CVFD requirements; add modflow 6.5.0 citation as "current version" in case we add full connectivity as an option to DIS/DISV

In models based on the Control Volume Finite Difference method, which is the method used in the MODFLOW family of hydrologic simulation codes \citep{modflow6gwf, langevin2024}, the loss of accuracy associated with the boundary-matching approach results from the use of a "two-point" formulation to calculate flows between adjacent grid cells. In the two-point formulation, the flow between two cells is proportional to the difference between the heads calculated in those two cells. For simulations of flow through isotropic porous media, the two-point flow formulation is most accurate when the grid satisfies the so-called ``CVFD requirement" \citep{modflow6gwf} that the straight-line connection between two cell centers must intersect the midpoint of the cell interface at a right angle \citep{narasimhan1976integrated}. However, boundary-matching grids violate the CVFD requirement at vertical interfaces between cells in the same model layer because the corresponding cell connections are not strictly horizontal.

@@

Figure \ref{fig:schem-conn-area-flux}a illustrates what we refer to as "layered" and "full" cell connectivity. The focus is on a group of four cells in a two-grid-layer, vertically offset MODFLOW 6 model of fully saturated flow through a permeable hydrogeologic layer (aquifer) with impermeable top and bottom boundaries. If the grid is represented using a layered grid type (DIS or DISV), the grid has what we call ``layered connectivity": a cell is hydraulically connected to each laterally adjacent cell in the same grid layer (regardless of whether or not the cell faces actually overlap), and with each overlying and underlying cell, as indicated by the blue lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a. In layered connectivity, two cells that have overlapping vertical faces but are in different grid layers (e.g., cells 1A and 2B) notably do not have a direct hydraulic connection. The addition of such connections between cells in different grid layers, as indicated by the red lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a, results in a grid with what we call ``full connectivity.''  Implementation of full connectivity in MODFLOW 6 requires that the vertically offset grid be represented using the fully unstructured (DISU) grid type.

>>>>

In models based on the Control Volume Finite Difference method, which is the method used in the MODFLOW family of hydrologic simulation codes \citep{modflow6gwf, langevin2024}, the loss of accuracy associated with the boundary-matching approach results from the use of a "two-point" formulation to calculate flows between adjacent grid cells. In the two-point formulation, the flow between two cells is proportional to the difference between the heads calculated in those two cells. For simulations of flow through isotropic porous media, the two-point flow formulation is most accurate when the grid satisfies the so-called ``CVFD requirement" \citep{modflowusg, modflow6gwf} that the straight-line connection between two cell centers must intersect the midpoint of the cell interface at a right angle \citep{narasimhan1976integrated}. However, boundary-matching grids violate the CVFD requirement at vertical interfaces between cells in the same model layer because the corresponding cell connections are not strictly horizontal.

@@

Figure \ref{fig:schem-conn-area-flux}a illustrates what we refer to as "layered" and "full" cell connectivity. The focus is on a group of four cells in a two-grid-layer, vertically offset MODFLOW 6 model of fully saturated flow through a permeable hydrogeologic layer (aquifer) with impermeable top and bottom boundaries. If the grid is represented using a layered grid type (DIS or DISV), the grid has what we call ``layered connectivity": a cell is hydraulically connected to each laterally adjacent cell in the same grid layer (regardless of whether or not the cell faces actually overlap), and with each overlying and underlying cell, as indicated by the blue lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a. In layered connectivity, two cells that have overlapping vertical faces but are in different grid layers (e.g., cells 1A and 2B) notably do not have a direct hydraulic connection. The addition of such connections between cells in different grid layers, as indicated by the red lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a, results in a grid with what we call ``full connectivity.''  Implementation of full connectivity in the current version of MODFLOW 6 \citep{modflow650software} requires that the vertically offset grid be represented using the fully unstructured (DISU) grid type.
